---
title: "Bicycle Sales Analysis"
subtitle: "Data Science for Business"
author: "Jake Weber"
date: "11/09/2021"
output: 
  html_document:
    theme: journal
    highlight: tango
    toc: TRUE
    number_sections: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

# Data Import

## Initiating Packages

```{r Installing Packages, eval=FALSE, include=TRUE}

pkgs_cran <- c(
    # File System
    "fs",         # working with the file system
    
    # Import
    "readr",     # reading excel files
    "writexl",    # saving data as excel files
    "odbc",       # connecting to databases
    "RSQLite",    # connecting to SQLite databases
    
    # Tidy, Transform, & Visualize
    "tidyverse",  # dplyr, ggplot2, tibble, tidyr, readr, purrr, stringr, forcats
    "lubridate",  # working with dates and times
    "tidyquant",  # used mainly for the ggplot plotting theme
    
    # Model
    "tidymodels", # installs broom, infer, recipes, rsample, & yardstick
    "umap",       # used for visualizing clusters
    
    # Other
    "devtools"    # used to install non-CRAN packages
)

install.packages("fs")       # Install single package
install.packages(pkgs_cran)  # Install many packages

```

```{r Inititating Packages}

library(tidyverse)
library(readxl)
library(writexl)
library(readr)
library(lubridate)
library(tidyquant)


```

## Reading the Data

```{r Reading in the Data}

# Set Up

options(scipen = 100)

getwd()

# DATA WRANGLING OVERVIEW ----

bikes_tbl <- read_excel("00_data/bike_sales/data_raw/bikes.xlsx")
orderlines_tbl <- read_excel("00_data/bike_sales/data_raw/orderlines.xlsx")
bike_orderlines_tbl <- read_rds("00_data/bike_sales/data_wrangled/bike_orderlines.rds")

bikeshops_tbl <- read_excel("00_data/bike_sales/data_raw/bikeshops.xlsx")

orderlines_tbl <- read_excel("00_data/bike_sales/data_raw/orderlines.xlsx")



```

## Glimpse of Bike Ordlines

```{r Glimse of Bike Orderlines}

#Glimpse gives a helpful overview showing the column names, data type, and the head of the data.

bike_orderlines_tbl %>% glimpse()  

```

# Selecting via select() 

```{r the Basic Select}

# Basic select
bike_orderlines_tbl %>%
    select(order_date, order_id, order_line)

bike_orderlines_tbl %>%
    select(1:3)

bike_orderlines_tbl %>%
    select(starts_with("order_"))

```

```{r Reducing the Columns Selected}

# Reduce columns

bike_orderlines_tbl %>%
    select(order_date, total_price, category_1, category_2)

bike_orderlines_tbl %>% 
    select(-state)

```

```{r Rearranging Columns}

# Rearange columns

bike_orderlines_tbl %>%
    select(bikeshop_name:state, everything())

```

```{r}

# pull() extracts a single column

bike_orderlines_tbl %>%
    # select(total_price) %>%
    pull(total_price) %>%
    mean()

#bike_orderlines_tbl %>% 
#    pull(model)

```

```{r}

# select_if

#?select_if

bike_orderlines_tbl %>%
    select_if(is.character) 

bike_orderlines_tbl %>%
    select_if(~ is.numeric(.))

bike_orderlines_tbl %>%
    select_if(~ !is.numeric(.))

```

# Sorting via arrange() 

```{r Data Arranged by Descending Price}

bikes_tbl %>%
    select(model, price) %>%
    arrange(desc(price)) %>%
    View()

```

# Filtering via filter() 

## Filter via formula 

```{r Formula Filtering}

bikes_tbl %>%
    select(model, price) %>%
    filter(price > mean(price))

bikes_tbl %>%
    select(model, price) %>%
    filter((price > 5000) | (price < 1000)) %>%
    arrange(desc(price)) %>%
    View()

bikes_tbl %>%
    select(model, price) %>%
    filter(price > 6000,
           model %>% str_detect("Supersix"))

# Filtering One or More Conditions Exactly Using == and %in%
bike_orderlines_tbl %>%
    filter(category_2 %in% c("Over Mountain", "Trail", "Endurance Road"))

bike_orderlines_tbl %>%
    filter(category_2 == "Over Mountain")

bike_orderlines_tbl %>%
    filter(category_2 != "Over Mountain")

bike_orderlines_tbl %>%
    filter(!(category_2 %in% c("Over Mountain", "Trail", "Endurance Road")))

```


## Filtering via slice()

```{r Slicing the Data}

bikes_tbl %>%
    arrange(desc(price)) %>%
    slice(1:5)

bikes_tbl %>%
    arrange(price) %>%
    slice(1:5)

bikes_tbl %>%
    arrange(desc(price)) %>%
    slice((nrow(.)-4):nrow(.))

bikes_tbl %>%
    arrange(desc(price)) %>%
    slice((93):97)

```


## Unique via distinct()

```{r Unique Values}

bike_orderlines_tbl %>%
    distinct(category_1)

bike_orderlines_tbl %>%
    distinct(category_1, category_2)

bike_orderlines_tbl %>%
    distinct(bikeshop_name, city, state)

```

# Creating via mutate() 

```{r}

# Adding column
bike_orderlines_prices <- bike_orderlines_tbl %>%
    select(order_date, model, quantity, price) %>%
    mutate(total_price = quantity * price)

bike_orderlines_prices

# Overwrite Column
bike_orderlines_prices %>%
    mutate(total_price = log(total_price)) 

# Transformations
bike_orderlines_prices %>%
    mutate(total_price_log = log(total_price)) %>%
    mutate(total_price_sqrt = total_price^0.5)

# Adding Flag
bike_orderlines_prices %>%
    mutate(is_supersix = model %>% str_to_lower() %>%
               str_detect("supersix")) %>%
    filter(is_supersix)

```

## Bins via ntile()

```{r Binning with ntile()}

bike_orderlines_prices %>%
    mutate(total_price_binned = ntile(total_price, 3)) %>% select(total_price_binned,everything())

```

## Flexibility via case_when()  

```{r case_when for flexibility}

# case_when() - more flexible binning

# Numeric to Categorical
bike_orderlines_prices %>%
    mutate(total_price_binned = ntile(total_price, 3)) %>%
    mutate(total_price_binned2 = case_when(
        total_price > quantile(total_price, 0.75) ~ "High",
        total_price > quantile(total_price, 0.25) ~ "Medium",
        TRUE ~ "Low"
    )) %>% select(total_price_binned2,total_price_binned,everything())

# Text to Categorical
bike_orderlines_prices %>%
    mutate(bike_type = case_when(
        model %>% str_to_lower() %>% str_detect("supersix") ~ "Supersix",
        model %>% str_to_lower() %>% str_detect("jekyll") ~ "Jekyll",
        TRUE ~ "Not Supersix or Jekyll"
    )) %>% select(bike_type, everything())

```


# Grouping & Summarizing

```{r}

# Basics 
bike_orderlines_tbl %>%
    summarise(
        revenue = sum(total_price)
    )

bike_orderlines_tbl %>%
    group_by(category_1) %>%
    summarise(revenue = sum(total_price))

bike_orderlines_tbl %>%
    group_by(category_1, category_2) %>%
    summarise(revenue = sum(total_price)) %>%
    ungroup() %>%
    arrange(desc(revenue))

bike_orderlines_tbl %>%
    group_by(category_1, category_2, frame_material) %>%
    summarise(revenue = sum(total_price)) %>%
    ungroup() %>%
    arrange(desc(revenue))

```

## Summary Functions

```{r Summary Functions}

bike_orderlines_tbl %>%
    group_by(category_1, category_2) %>%
    summarize(
        count = n(),
        avg   = mean(total_price),
        med   = median(total_price),
        sd    = sd(total_price),
        min   = min(total_price),
        max   = max(total_price)
    ) %>%
    ungroup() %>%
    arrange(desc(count))

```

## Detecting NAs

```{r Summarize all}

# summarize_all() - detect missing values

bike_orderlines_missing <- bike_orderlines_tbl %>%
    mutate(total_price = c(rep(NA, 4), total_price[5:nrow(.)]))

bike_orderlines_missing %>%
    summarise_all(~ sum(is.na(.)))

bike_orderlines_missing %>%
    summarise_all(~ sum(is.na(.)) / length(.))

bike_orderlines_missing %>%
    filter(!is.na(total_price))

```

# rename() and set_names()

## rename: One at a time

```{r rename()}

bikeshop_revenue_tbl <- bike_orderlines_tbl %>%
    select(bikeshop_name, category_1, total_price) %>%
    
    group_by(bikeshop_name, category_1) %>%
    summarize(sales = sum(total_price)) %>%
    ungroup() %>%
    
    arrange(desc(sales))

bikeshop_revenue_tbl %>%
    rename(
        `Bikeshop Name` = bikeshop_name,
        `Primary Category` = category_1,
        Sales = sales
    )

```


## set_names: All at once

```{r set_names()}

bikeshop_revenue_tbl %>%
    set_names(c("Bikeshop Name", "Primary Category", "Sales"))

bikeshop_revenue_tbl %>%
    set_names(names(.) %>% str_replace("_", " ") %>% str_to_title())


```

# Pivoting Data

## spread(): Long to Wide

```{r}
bikeshop_revenue_formatted_tbl <- bikeshop_revenue_tbl %>%
    
    spread(key = category_1, value = sales) %>%
    arrange(desc(Mountain)) %>%
    rename(`Bikeshop Name` = bikeshop_name) %>%
    mutate(
        Mountain = scales::dollar(Mountain),
        Road     = scales::dollar(Road)
    )

bikeshop_revenue_formatted_tbl

```


## gather(): Wide to Long

```{r}
bikeshop_revenue_formatted_tbl %>%
    gather(key = "category_1", value = "sales", Mountain, Road) %>%
    mutate(sales = sales %>% str_remove_all("\\$|,") %>% as.double()) %>%
    arrange(desc(sales))
```


# Joining Data

```{r}
#?left_join

orderlines_tbl %>%
    left_join(y = bikes_tbl, by = c("product.id" = "bike.id"))
```


# Column binding

## bind_cols()

```{r}
bike_orderlines_tbl %>%
    select(-contains("order")) %>%
    bind_cols(bike_orderlines_tbl %>% select(order_id))
```


## bind_rows()

```{r}
train_tbl <- bike_orderlines_tbl %>%
    slice(1:(nrow(.)/2))

train_tbl

test_tbl <- bike_orderlines_tbl %>%
    slice((nrow(.)/2 + 1):nrow(.))

test_tbl

train_tbl %>%
    bind_rows(test_tbl)

```


# Separate & Unite

```{r}
bike_orderlines_tbl %>%
    select(order_date) %>%
    mutate(order_date = as.character(order_date)) %>%
    
    # separate
    separate(col = order_date, into = c("year", "month", "day"), sep = "-", remove = FALSE) %>%
    
    mutate(
        year  = as.numeric(year),
        month = as.numeric(month),
        day   = as.numeric(day)
    ) %>%
    
    # unite
    unite(order_date_united, year, month, day, sep = "-", remove = FALSE) %>%
    mutate(order_date_united = as.Date(order_date_united))
```

# lubridate()

```{r}
# 1.0 Date & Lubridate Basics ----

# 1.1 Character vs Date/Datetime

order_date_tbl <- bike_orderlines_tbl %>%
    select(order_date)

order_date_tbl %>%
    pull(order_date) %>%
    class()


# 1.2 Date Classes

order_date_tbl %>%
    mutate(order_date_chr = as.character(order_date)) %>%
    mutate(order_date_chr2 = order_date_chr %>% str_c(" 00:00:00")) %>%
    
    mutate(order_date_date = order_date_chr %>% ymd()) %>%
    mutate(order_date_dttm = order_date_chr2 %>% ymd_hms())





# 1.3 Lubridate Functions

# Conversion

"06/01/18" %>% mdy() %>% class()

"06/01/18 12:30:15" %>% mdy_hms() %>% class()

"January 1, 1985" %>% mdy()



# Extractor

"2011-01-01" %>% ymd() %>% year()

"2011-01-01" %>% ymd() %>% month(label = TRUE, abbr = FALSE) 

"2011-01-01" %>% ymd() %>% wday(label = TRUE, abbr = FALSE)

"2011-01-01" %>% ymd() %>% day()

# Helpers

now()

today()

# Periods & Durations - Add/subract time to/from a date

today() + days(12)

today() + ddays(12)

today() + years(4)  # Period

today() + dyears(4) # Duration

# Intervals - Calculate time-based distance 

i <- interval(today(), today() + ddays(12)) 

i / ddays(1) # interval / ddays = how many days in interval

i / dminutes(1) # interval / dminutes = how many minutes in the interval

order_date_tbl %>%
    mutate(today = today()) %>%
    mutate(diff_days = interval(order_date, today) / ddays(1))


# 2.0 Time-Based Data Grouping ----

bike_sales_y_tbl <- bike_orderlines_tbl %>%
    select(order_date, total_price) %>%
    
    # lubridate
    mutate(order_date = ymd(order_date)) %>%
    mutate(year = year(order_date)) %>%
    
    # group_by + summarize
    group_by(year) %>%
    summarize(sales = sum(total_price)) %>%
    ungroup()

bike_sales_y_tbl

bike_sales_m_tbl <- bike_orderlines_tbl %>%
    select(order_date, total_price) %>%
    
    # lubridate
    mutate(order_date = ymd(order_date)) %>%
    mutate(
        year  = year(order_date),
        month = month(order_date, label = TRUE, abbr = TRUE)
    ) %>%
    
    # groupby + summarize
    group_by(year, month) %>%
    summarize(sales = sum(total_price)) %>%
    ungroup()

bike_sales_m_tbl

# Floor Date

bike_orderlines_tbl %>%
    select(order_date, total_price) %>%
    
    # lubridate
    mutate(order_date = ymd(order_date)) %>%
    mutate(year_month = floor_date(order_date, unit = "month")) %>%
    
    # group_by + summarize
    group_by(year_month) %>%
    summarize(sales = sum(total_price))

# 3.0 Measuring Change ----

# 3.1 Difference from most recent observation ----

bike_sales_y_tbl %>%
    mutate(sales_lag_1 = lag(sales, n = 1)) %>%
    
    # Handle NA
    mutate(sales_lag_1 = case_when(
        is.na(sales_lag_1) ~ sales,
        TRUE ~ sales_lag_1
    )) %>%
    
    # Diff's & Pct Diffs
    mutate(diff_1 = sales - sales_lag_1) %>%
    mutate(pct_diff_1 = diff_1 / sales_lag_1) %>%
    mutate(pct_diff_1_chr = scales::percent(pct_diff_1))

calculate_pct_diff <- function(data) {
    
    data %>% 
        mutate(sales_lag_1 = lag(sales, n = 1)) %>%
        
        # Handle NA
        mutate(sales_lag_1 = case_when(
            is.na(sales_lag_1) ~ sales,
            TRUE ~ sales_lag_1
        )) %>%
        
        # Diff's & Pct Diffs
        mutate(diff_1 = sales - sales_lag_1) %>%
        mutate(pct_diff_1 = diff_1 / sales_lag_1) %>%
        mutate(pct_diff_1_chr = scales::percent(pct_diff_1))
    
}

bike_sales_m_tbl %>%
    calculate_pct_diff()



# 3.2 Difference from first observation ----

bike_sales_y_tbl %>%
    mutate(sales_2011 = first(sales)) %>%
    mutate(diff_2011  = sales - sales_2011) %>%
    mutate(pct_diff_2011 = diff_2011 / sales_2011) %>%
    mutate(pct_diff_2011_chr = scales::percent(pct_diff_2011))

bike_sales_m_tbl %>%
    
    group_by(year) %>%
    
    mutate(sales_jan = first(sales)) %>% 
    mutate(
        diff_jan         = sales - sales_jan,
        pct_diff_jan     = diff_jan / sales_jan,
        pct_diff_jan_chr = scales::percent(pct_diff_jan)
    ) 



# 4.0 Cumulative Calculations ----

bike_sales_y_tbl %>%
    mutate(cumulative_sales = cumsum(sales)) %>%
    mutate(cumulative_sales_pct = cumulative_sales / sum(sales)) %>%
    mutate(cumulative_sales_pct_chr = cumulative_sales_pct %>% scales::percent())

bike_sales_m_tbl %>%
    
    group_by(year) %>%
    
    mutate(cumulative_sales = cumsum(sales)) %>%
    mutate(cumulative_sales_pct = cumulative_sales / sum(sales)) %>%
    mutate(cumulative_sales_pct_chr = scales::percent(cumulative_sales_pct))


# 5.0 Rolling Calculations ----

bike_sales_m_tbl %>%
    
    mutate(roll_mean_3 = rollmean(sales, k = 3, na.pad = TRUE, align = "right", fill = NA)) %>%
    
    mutate(roll_mean_6 = rollmean(sales, k = 6, na.pad = TRUE, align = "right", fill = NA))


# 6.0 Filtering Date Ranges ---- 

bike_orderlines_tbl %>%
    
    mutate(order_date = ymd(order_date)) %>%
    
    filter(order_date %>% between(left = ymd("2012-01-01"), right = ymd("2013-12-31"))) 


bike_orderlines_tbl %>%
    
    mutate(order_date = ymd(order_date)) %>%
    
    filter(year(order_date) %in% c(2012, 2013))
```


# stringr()

```{r}

# TEXT MANIPULATION ----

# 1.0 Basics ----

# 1.1 Detection: Used with filter() ----

# Vector
c("Supersix Evo Black Inc.", "Supersix Evo Hi-Mod Team") %>%
    str_detect(pattern = "Supersix")

# Tibble
bikes_tbl %>%
    select(model) %>%
    mutate(supersix = model %>% str_detect("Supersix") %>% as.numeric()) %>% 
    mutate(black    = model %>% str_detect("Black") %>% as.numeric())


# 1.2 Case & Concatenation ----


# Case
bikeshop_name <- "Ithaca Mountain Climbers"

str_to_upper(bikeshop_name)
str_to_lower(bikeshop_name)
str_to_title(bikeshop_name)

# Concatenation

# Vector
order_id <- 1
order_line <- 1

str_c("Order Line: ", order_id, ".", order_line, 
      " sent to Customer: ", bikeshop_name,
      sep = "")

str_glue("Order Line: {order_id}.{order_line} sent to Customer: {str_to_upper(bikeshop_name)}")


# Tibble
bike_orderlines_tbl %>%
    select(bikeshop_name, order_id, order_line) %>%
    mutate(purchase_statement = str_glue(
        "Order Line: {order_id}.{order_line} sent to Customer: {str_to_upper(bikeshop_name)}"
    ) %>% as.character())

# 1.3 Separating Text: See tidyr::separate() ----

# Vector
c("Road - Elite Road - Carbon", "Road - Elite Road") %>% str_split(pattern = " - ", simplify = TRUE)

# Tibble
bikes_tbl %>%
    select(description) %>%
    separate(col    = description, 
             into   = c("category_1", "category_2", "frame_material"), 
             sep    = " - ",
             remove = FALSE)


# 1.4 Trimming Text ----

" text with space   " %>% str_trim(side = "left")

# 1.5 Replacement: Used with mutate() [and optionally case_when()] ----

# Vector
c("CAAD12", "CAAD", "CAAD8") %>% str_replace(pattern = "[0-9]", replacement = "")

c("CAAD12", "CAAD", "CAAD8") %>% str_replace_all(pattern = "[0-9]", replacement = "")

# Tibble
bikes_tbl %>%
    select(model) %>%
    mutate(model_num_removed = model %>% str_replace_all("[0-9]", "") %>% str_trim()) 



# 1.6 Formatting Numbers ----

# values
value <- 1e6

(value / 1e6) %>% scales::number(prefix = "$", suffix = "M")

value %>% scales::number(prefix = "$", big.mark = ",")

value %>% scales::dollar(scale = 1/1e6, suffix = "M")

# percents
pct <- 0.15

pct %>% scales::number(scale = 100, suffix = "%")

pct %>% scales::percent()


# 1.7 Formatting Column Names ----

# Replacing text in column names

bike_orderlines_tbl %>%
    set_names(names(.) %>% str_replace("_", ".") %>% str_to_upper())

# Appending text to column names
bike_orderlines_tbl %>%
    set_names(str_glue("{names(.)}_bike"))

# Appending text to specific column names
bike_orderlines_colnames_tbl <- bike_orderlines_tbl %>%
    rename_at(.vars = vars(model:frame_material), 
              .funs = ~ str_c("prod_", .)) %>%
    rename_at(vars(bikeshop_name:state),
              ~ str_c("cust_", .)) 

bike_orderlines_colnames_tbl %>%
    select(contains("cust_"), total_price)

# 2.0 Feature Engineering with Text -----
# Investigating "model" and extracting well-formatted features

bikes_tbl %>%
    
    select(model) %>%
    
    # Fix typo
    mutate(model = case_when(
        model == "CAAD Disc Ultegra" ~ "CAAD12 Disc Ultegra",
        model == "Syapse Carbon Tiagra" ~ "Synapse Carbon Tiagra",
        model == "Supersix Evo Hi-Mod Utegra" ~ "Supersix Evo Hi-Mod Ultegra",
        TRUE ~ model
    )) %>%
    
    # separate using spaces
    separate(col     = model, 
             into    = str_c("model_", 1:7), 
             sep     = " ", 
             remove  = FALSE, 
             fill    = "right") %>%
    
    # creating a "base" feature
    mutate(model_base = case_when(
        
        # Fix Supersix Evo
        str_detect(str_to_lower(model_1), "supersix") ~ str_c(model_1, model_2, sep = " "),
        
        # Fix Fat CAAD bikes
        str_detect(str_to_lower(model_1), "fat") ~ str_c(model_1, model_2, sep = " "),
        
        # Fix Beast of the East
        str_detect(str_to_lower(model_1), "beast") ~ str_c(model_1, model_2, model_3, model_4, sep = " "),
        
        # Fix Bad Habit
        str_detect(str_to_lower(model_1), "bad") ~ str_c(model_1, model_2, sep = " "),
        
        # Fix Scalpel 29
        str_detect(str_to_lower(model_2), "29") ~ str_c(model_1, model_2, sep = " "),
        
        # catch all
        TRUE ~ model_1)
    ) %>%
    
    # Get "tier" feature
    mutate(model_tier = model %>% str_replace(model_base, replacement = "") %>% str_trim()) %>%
    
    # Remove unnecessary columns
    select(-matches("[0-9]")) %>%
    
    # Create Flags
    mutate(
        black     = model_tier %>% str_to_lower() %>% str_detect("black") %>% as.numeric(),
        hi_mod    = model_tier %>% str_to_lower() %>% str_detect("hi-mod") %>% as.numeric(),
        team      = model_tier %>% str_to_lower() %>% str_detect("team") %>% as.numeric(),
        red       = model_tier %>% str_to_lower() %>% str_detect("red") %>% as.numeric(),
        ultegra   = model_tier %>% str_to_lower() %>% str_detect("ultegra") %>% as.numeric(),
        dura_ace  = model_tier %>% str_to_lower() %>% str_detect("dura ace") %>% as.numeric(),
        disc      = model_tier %>% str_to_lower() %>% str_detect("disc") %>% as.numeric()
    ) %>%
    
    View()


```

# forcats()

```{r}
# 1.0 Factor Basics ----

# What is a Factor?
# A way of managing categorical data

# Why do we want factors? 
# 1. Can group numeric values into bin (think price = low, medium, high)
# 2. Can reorder categories for visualization (fct_reorder)
# 3. Can manipulate categories much eaiser (fct_lump)
# 4. Machine learning and modeling algorithms may require factor data type for categorical data. 

# 2.0 Motivating Example -----

# Manipulation
sales_by_cat_2_tbl <- bike_orderlines_tbl %>%
    
    select(category_2, total_price) %>%
    
    group_by(category_2) %>%
    summarize(sales = sum(total_price)) %>%
    ungroup() %>%
    
    arrange(desc(sales)) %>%
    
    mutate(category_2 = category_2 %>% as_factor() %>% fct_rev())

# Plotting
sales_by_cat_2_tbl %>%
    ggplot(aes(x = sales, y = category_2)) +
    geom_point(size = 5, color = "#2c3e50") +
    labs(title = "Sales By Category 2") +
    scale_x_continuous(labels = scales::dollar_format()) +
    theme_tq() +
    expand_limits(x = 0)

plot_sales <- function(data) {
    data %>%
        ggplot(aes(x = sales, y = category_2)) +
        geom_point(size = 5, color = "#2c3e50") +
        labs(title = "Sales By Category 2") +
        scale_x_continuous(labels = scales::dollar_format()) +
        theme_tq() +
        expand_limits(x = 0)
}

sales_by_cat_2_tbl %>%
    plot_sales()



# 3.0 Forcats Basics ----


# 3.1 Inspecting Factors ----

# Vector
sales_by_cat_2_tbl %>% pull(category_2) %>% levels()

sales_by_cat_2_tbl %>% pull(category_2) %>% as.numeric()

# Tibble

sales_by_cat_2_tbl %>%
    mutate(category_2 = category_2 %>% fct_rev() %>% fct_rev()) %>%
    mutate(
        label = category_2 %>% as.character(),
        value = category_2 %>% as.numeric()
    )

# 3.2 Creating Factors: as_factor() vs as.factor() ----

sales_by_cat_2_tbl %>%
    mutate(
        category_2           = as.character(category_2),
        category_2_as_factor = as_factor(category_2) %>% as.numeric(),
        category_2_as.factor = as.factor(category_2) %>% as.numeric()
    )



# 3.3 Reordering Factors: fct_reorder() and fct_rev() ----

sales_by_cat_2_tbl %>%
    arrange(desc(sales)) %>%
    mutate(sales_negative = -sales) %>%
    mutate(
        category_2 = category_2 %>% fct_reorder(sales_negative),
        values     = category_2 %>% as.numeric()) %>%
    
    plot_sales()





# 3.4 Time-Based Reordering: fct_reorder2() ----


sales_by_cat_2_q_tbl <- bike_orderlines_tbl %>%
    
    mutate(order_date = order_date %>% floor_date("quarter") %>% ymd()) %>%
    
    group_by(category_2, order_date) %>%
    summarise(sales = sum(total_price)) %>%
    ungroup()

sales_by_cat_2_q_tbl


sales_by_cat_2_q_tbl %>%
    
    mutate(category_2 = category_2 %>% fct_reorder2(order_date, sales)) %>%
    
    ggplot(aes(x = order_date, y = sales, color = category_2)) +
    geom_point() +
    geom_line() +
    facet_wrap(~ category_2) +
    
    theme_tq() +
    scale_color_tq() +
    scale_y_continuous(labels = scales::dollar_format(scale = 1e-6, suffix = "M"))


# 3.5 Creating "Other" Category - fct_lump() & fct_relevel() ----

sales_by_cat_2_tbl %>%
    
    mutate(category_2 = category_2 %>% fct_lump(n = 6, 
                                                w = sales, 
                                                other_level = "All Other Bike Categories")) %>%
    
    group_by(category_2) %>%
    summarize(sales = sum(sales)) %>%
    
    mutate(category_2 = category_2 %>% fct_relevel("All Other Bike Categories", after = 0)) %>%
    
    plot_sales()

```


# Visualization

```{r}
left_join(orderlines_tbl, bikes_tbl, by = c("product.id" = "bike.id"))

bike_orderlines_joined_tbl <- orderlines_tbl %>%
    left_join(bikes_tbl, by = c("product.id" = "bike.id")) %>%
    left_join(bikeshops_tbl, by = c("customer.id" = "bikeshop.id"))

bike_orderlines_joined_tbl

bike_orderlines_joined_tbl %>% glimpse()

# 5.0 Wrangling Data ----

bike_orderlines_wrangled_tbl <- bike_orderlines_joined_tbl %>%
    
    # Separate description into category.1, category.2, and frame.material
    separate(description,
             into = c("category.1", "category.2", "frame.material"),
             sep = " - ",
             remove = TRUE) %>%
    
    # Separate location into city and state
    separate(location,
             into = c("city", "state"),
             sep  = ", ",
             remove = FALSE) %>%
    
    # price extended
    mutate(total.price = price * quantity) %>%
    
    # Reorganize
    select(-location) %>%
    select(-ends_with(".id")) %>%
    
    bind_cols(bike_orderlines_joined_tbl %>% select(order.id)) %>%
    
    # Reorder columns
    select(contains("date"), contains("id"), contains("order"),
           quantity, price, total.price,
           everything()) %>%
    
    # Renaming columns
    rename(order_date = order.date) %>%
    set_names(names(.) %>% str_replace_all("\\.", "_")) 

bike_orderlines_wrangled_tbl %>% glimpse()



# 6.0 Business Insights ----


# 6.1 Sales by Year ----

# Step 1 - Manipulate

sales_by_year_tbl <- bike_orderlines_wrangled_tbl %>%
    
    # Selecting columns to focus on and adding a year column
    select(order_date, total_price) %>%
    mutate(year = year(order_date)) %>%
    
    # Grouping by year, and summarizing sales
    group_by(year) %>%
    summarize(sales = sum(total_price)) %>%
    ungroup() %>%
    
    # $ Format Text
    mutate(sales_text = scales::dollar(sales))
    
sales_by_year_tbl


# Step 2 - Visualize

sales_by_year_tbl %>%
    
    # Setup canvas with year (x-axis) and sales (y-axis)
    ggplot(aes(x = year, y = sales)) +
    
    # Geometries
    geom_col(fill = "#2c3e50") +
    geom_label(aes(label = sales_text)) +
    geom_smooth(method = "lm", se = FALSE) +
    
    # Formatting
    theme_tq() +
    scale_y_continuous(labels = scales::dollar) +
    labs(
        title = "Revenue by Year",
        subtitle = "Upward trend",
        x = "",
        y = "Revenue"
    )

# 6.2 Sales by Year and Category 2 ----


# Step 1 - Manipulate

sales_by_year_cat_2_tbl <- bike_orderlines_wrangled_tbl %>%
    
    # Selecting columns and add a year
    select(order_date, total_price, category_2) %>%
    mutate(year = year(order_date)) %>%
    
    # Groupby and Summarize year and category 2
    group_by(year, category_2) %>%
    summarise(sales = sum(total_price)) %>%
    ungroup() %>%
    
    # Format $ Text
    mutate(sales_text = scales::dollar(sales))


sales_by_year_cat_2_tbl

# Step 2 - Visualize

sales_by_year_cat_2_tbl %>%
    
    # Set up x, y, fill 
    ggplot(aes(x = year, y = sales, fill = category_2)) +
    
    # Geometries
    geom_col() +
    geom_smooth(method = "lm", se = FALSE) +
    
    # Facet
    facet_wrap(~ category_2, ncol = 3, scales = "free_y") +
    
    # Formatting
    theme_tq() +
    scale_fill_tq() +
    scale_y_continuous(labels = scales::dollar) +
    labs(
        title = "Revenue by Year and Category 2",
        subtitle = "Each product category has an upward trend",
        x = "",
        y = "Revenue",
        fill = "Product Secondary Category"
    )


# 7.0 Writing Files ----

fs::dir_create("00_data/bike_sales/data_wrangled_student")

# 7.1 Excel ----

bike_orderlines_wrangled_tbl %>%
    write_xlsx("00_data/bike_sales/data_wrangled_student/bike_orderlines.xlsx")


# 7.2 CSV ----

bike_orderlines_wrangled_tbl %>%
    write_csv("00_data/bike_sales/data_wrangled_student/bike_orderlines.csv")


# 7.3 RDS ----

bike_orderlines_wrangled_tbl %>%
    write_rds("00_data/bike_sales/data_wrangled_student/bike_orderlines.rds")


```

## Formating

```{r}

# Data Manipulation

sales_by_year_category_2_tbl <- bike_orderlines_tbl %>%
    select(order_date, category_2, total_price) %>%
    
    mutate(order_date = ymd(order_date)) %>%
    mutate(year = year(order_date)) %>%
    
    group_by(category_2, year) %>%
    summarize(revenue = sum(total_price)) %>%
    ungroup() %>%
    
    mutate(category_2 = fct_reorder2(category_2, year, revenue))

sales_by_year_category_2_tbl

sales_by_year_category_2_tbl %>%
    mutate(category_2_num = as.numeric(category_2)) %>%
    arrange(category_2_num)

# 1.0 Working with Colors ----

# 1.1 Color Conversion ----

# Named Colors
colors()

sales_by_year_category_2_tbl %>%
    
    ggplot(aes(year, revenue)) +
    
    geom_col(fill = viridisLite::viridis(n = 20)[10])

# To RGB
col2rgb("slateblue")

col2rgb("#2C3E50")

# To HEX
rgb(44, 62, 80, maxColorValue = 255)
 

# 1.2 Color Palettes ----

# tidyquant
tidyquant::palette_light()

palette_light()[2] %>% col2rgb()

# Brewer
RColorBrewer::display.brewer.all()
RColorBrewer::brewer.pal.info
RColorBrewer::brewer.pal(n = 8, name = "Blues")

# Viridis
viridisLite::viridis(n = 20)



# 2.0 Aesthetic Mappings ----

# 2.1 Color  -----
# - Used with line and points, Outlines of rectangular objects
sales_by_year_category_2_tbl %>%
    
    ggplot(aes(year, revenue, color = category_2)) +
    geom_line(size = 1) +
    geom_point(color = "dodgerblue", size = 5)


# 2.2 Fill  -----
# - Used with fill of rectangular objects 

sales_by_year_category_2_tbl %>%
    
    ggplot(aes(year, revenue)) +
    geom_col(aes(fill = category_2))

# 2.3 Size ----
# - Used with points

sales_by_year_category_2_tbl %>%
    
    ggplot(aes(year, revenue, size = revenue)) +
    geom_line(aes(color = category_2), size = 1) +
    geom_point()+
    facet_wrap(~ category_2)



# 3.0 Faceting ----
# - Great way to tease out variation by category

# Goal: Sales annual sales by category 2

sales_by_year_category_2_tbl %>%
    
    ggplot(aes(year, revenue, color = category_2)) +
    geom_line(color = "black") +
    geom_smooth(method = "lm", se = FALSE) +
    
    facet_wrap(~ category_2, ncol = 3, scales = "free_y") +
    
    expand_limits(y = 0)


# 4.0 Position Adjustments (Stack & Dodge) ----

# Stacked Bars & Side-By-Side Bars

sales_by_year_category_2_tbl %>%
    
    ggplot(aes(year, revenue, fill = category_2)) +
    # geom_col(position = "stack")
    # geom_col(position = "dodge")
    geom_col(position = position_dodge(width = 0.9), color = "white")

# Stacked Area

sales_by_year_category_2_tbl %>%
    
    ggplot(aes(year, revenue, fill = category_2)) +
    geom_area(color = "black")




# 5.0 Scales (Colors, Fills, Axis) ----

# 5.1 Plot Starting Points ----
# - Continuous (e.g. Revenue): Changes color via gradient palette
# - Categorical (e.g. ): Changes color via discrete palette

# Plot 1: Faceted Plot, Color = Continuous Scale
g_facet_continuous <- sales_by_year_category_2_tbl %>%
    
    ggplot(aes(year, revenue, color = revenue)) +
    geom_line(size = 1) +
    geom_point(size = 3) +
    
    facet_wrap(~ category_2, scales = "free_y") +
    expand_limits(y = 0) +
    
    theme_minimal()

g_facet_continuous
    

# Plot 2: Faceted Plot, Color = Discrete Scale
g_facet_discrete <- sales_by_year_category_2_tbl %>%
    
    ggplot(aes(year, revenue, color = category_2)) +
    geom_line(size = 1) +
    geom_point(size = 3) +
    
    facet_wrap(~ category_2, scales = "free_y") +
    expand_limits(y = 0) +
    
    theme_minimal()

g_facet_discrete

# Plot 3: Stacked Area Plot
g_area_discrete <- sales_by_year_category_2_tbl %>%
    
    ggplot(aes(year, revenue, fill = category_2)) +
    geom_area(color = "black") +
    
    theme_minimal()

g_area_discrete

# 5.2 Scale Colors & Fills ----
# - Awesome way to show variation by groups (discrete) and by values (continuous)

# Color by Revenue (Continuous Scale)

g_facet_continuous +
    
    # scale_color_continuous(
    #     low   = "black",
    #     high  = "cornflowerblue" 
    # )
    scale_color_viridis_c(direction = 1)
    


# Color by Category 2 (Discrete Scale)
RColorBrewer::display.brewer.all()
RColorBrewer::brewer.pal.info
RColorBrewer::brewer.pal(n = 8, name = "Blues")

g_facet_discrete +
    scale_color_brewer(palette = "Set3") +
    theme_dark()

g_facet_discrete +
    scale_color_tq(theme = "dark") +
    theme_dark()

g_facet_discrete +
    scale_color_viridis_d(option = "D") +
    theme_dark()


# Fill by Category 2

g_area_discrete +
    scale_fill_brewer(palette = "Set3")

g_area_discrete +
    scale_fill_tq()

g_area_discrete +
    scale_fill_viridis_d(direction = -1)


# 5.3 Axis Scales ----

sales_by_year_category_2_tbl

g_facet_continuous +
    scale_x_continuous(breaks = seq(2011, 2015, by = 2)) +
    scale_y_continuous(labels = scales::dollar_format(scale = 1e-6, suffix = "M"))



# 6.0 Labels ----

g_facet_continuous +
    
    scale_x_continuous(breaks = seq(2011, 2015, by = 2)) +
    scale_y_continuous(labels = scales::dollar_format(scale = 1e-6, suffix = "M")) +
    
    geom_smooth(method = "lm", se = FALSE) +
    
    scale_color_viridis_c() +
    theme_dark() +
    
    labs(
        title = "Bike Sales",
        subtitle = "Sales are trending up",
        caption = "5-year sales trends\ncomes from our ERP Database",
        x = "Year",
        y = "Revenue ($M)",
        color = "Revenue"
    )


# 7.0 Themes  ----

g_facet_continuous +
    
    theme_light() +
    
    theme(
        axis.text.x = element_text(
            angle = 45, 
            hjust = 1
        ),
        strip.background = element_rect(
            color = "black",
            fill  = "cornflowerblue",
            size  = 1
        ),
        strip.text = element_text(
            face  = "bold",
            color = "white"
        )
    )


# 8.0 Putting It All Together ----

# Brewer
RColorBrewer::display.brewer.all()
RColorBrewer::brewer.pal.info
RColorBrewer::brewer.pal(n = 9, name = "Blues")

sales_by_year_category_2_tbl %>%
    
    ggplot(aes(year, revenue, fill = category_2)) +
    
    geom_area(color = "black") +
    
    # Scales
    scale_fill_viridis_d(direction = 1) +
    scale_y_continuous(labels = scales::dollar_format()) +
    
    # Labels
    labs(
        title = "Sales Over Year by Category 2",
        subtitle = "Sales Trending Upward",
        x = "",
        y = "Revenue ($M)",
        fill = "2nd Category",
        caption = "Bike sales trends look strong heading into 2016"
    ) +
    
    # Theme
    theme_light() +
    theme(
        title = element_text(face = "bold")
    )



```


























